Lattice model

### 2017-04-01 oopsla17/lattice-model.tex
type setting fix
Fix projection signature

\begin{defn}{(\zinf)}
  We define the set \zinf{} as $\mathbb{Z}$ extended with $\infty$ and
  $-\infty$. For any $n$ in \zinf{}, we have $-\infty \leq n \leq \infty$. The
  resulting set is a total order with top and bottom elements.
\end{defn}

\begin{defn}{(\emph{Offset space})}
  \emph{Offset space} is a subset of \zinf{} members. It is used to capture
  offsets from induction variables.
\end{defn}

\begin{defn}{(\emph{\zinf{} interval})}
  A \emph{\zinf{} interval} is of the form \interv{a}{b}{c}, where $a$ and $b$
  are drawn from \zinf{} with $a \leq 0 \leq b$, $a$ is $-\infty$ \emph{iff} $b$
  is $\infty$, and $c$ is drawn from $\{ \mathsf{true}, \mathsf{false} \}$. We
  define it as follows:
%
  \begin{equation*}
    \interv{a}{b}{c} \triangleq
      \{ n \;|\; a \leq n \leq b \wedge (\neg c \implies n \neq 0) \}
  \end{equation*}

@@

\begin{defn}{(\emph{Offsets vector})}
  An $N$ dimensional \emph{offsets vector} is a Cartesian product of offset
  spaces denoted with $\textit{offs}_1 \times \cdots \times \textit{offs}_N$.
  The function $\pi_i : \zinf{}^{\mkern-19mu{N}} \to \zinf{}$ retrieves the
  $i^{th}$ set of offsets.

  The set of $N$ dimensional offsets vectors is
  $\mathcal{P}{(\zinf{}^{\mkern-19mu{N}})}$ and they trivially form a
  distributive lattice under the superset relation.
\end{defn}

\begin{defn}{(\emph{Point offsets vector})}
  A \emph{point offsets vector}, \textit{poff}, is a an offsets vector such that
  for all indices $i$ $\pi_i(\textit{poff})$, we either have a singleton set or
  the infinite interval $\interv{-\infty}{\infty}{\mathsf{true}}$.
\end{defn}

\begin{lem}{}\label{lem:vector-intersect}
  Intersection distributes over offsets vectors. If $\mathit{off}$ and
  $\mathit{off'}$ are $N$ dimensional ofssets vectors, then
%
  \begin{equation*}
    \mathit{off} \cap \mathit{off'} =
      \prod_{i = 1}^{N} \pi_i(\mathit{off}) \cap \pi_i(\mathit{off'})
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    \mathit{off} \; \cap \; \mathit{off'} &
    = \{x \mid \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off}) \}
      \;\cap\;
      \{x \mid \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off'}) \} \\
    & = \{x \mid \bigwedge_{1 \leq i \leq N }
      (x_i \in \pi_i(\mathit{off}) \wedge x_i \in \pi_i(\mathit{off'})) \} \\
    & = \prod_{i = 1}^{N} \pi_i(\mathit{off}) \cap \pi_i(\mathit{off'})
  \end{align*}
\end{proof}

\begin{lem}{}\label{lem:vector-union}
  Let $\mathit{off}$ and $\mathit{off'}$ be two $N$ dimensional offsets vectors
  such that $\pi_i(\mathit{off})$ is equal to $\pi_i(\mathit{off'})$ for all $1
  \leq i \leq N$ except possibly at $k$, then we have the following:
%
  \begin{equation*}
    \mathit{off} \; \cup \; \mathit{off'}
    =
    \pi_1(\mathit{off}) \times \cdots \times
    (\pi_k(\mathit{off}) \cup \pi_k(\mathit{off'})) \times \cdots \times
    \pi_N(\mathit{off})
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    \mathit{off} \; \cup \; \mathit{off'} &
    = \{x \mid
      \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off}) \}
      \;\cup\;
      \{x \mid
          \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off'}) \} \\
    & = \{x \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}}
            x_i \in \pi_i(\mathit{off}) \wedge x_k \in \pi_k(\mathit{off}) \vee
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in
            \pi_i(\mathit{off'}) \wedge x_k \in \pi_k(\mathit{off'})
        \} \\
    & = \{x \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in
            \pi_i(\mathit{off}) \wedge
            x_k \in \pi_k(\mathit{off}) \cup \pi_k(\mathit{off'})
        \} \\
        & = \pi_1(\mathit{off}) \times \cdots \times
        (\pi_k(\mathit{off}) \cup \pi_k(\mathit{off'})) \times \cdots \times
        \pi_N(\mathit{off})
  \end{align*}
\end{proof}

\begin{defn}{(\emph{Vector interval})}
  A vector interval is an $N$ dimensional object which has \zinf{} intervals at
  each dimension. A vector interval is also an ofssets vector.

  The set of all $N$ dimensional vector intervals is $\textit{Interval}^N$.
\end{defn}

\begin{defn}{(\emph{Region})}
  \region{N} is the set of $N$ dimensional vector intervals defined as
  the smallest set satisfying the following:
%
  \begin{enumerate}
    \item If $R$ is in $\textit{Interval}^N$, then $R$ is in \region{N}.
    \item If $R$ and $S$ are in \region{N}, then so are $R \cap S$ and
      $R \cup S$.
  \end{enumerate}
\end{defn}

@@

Here helper function $\textit{peel} : \mathsf{Approx} a \to a$ removes the
approximation label.

@@

\begin{defn}{(Extended origin)}
  In the $N$ dimensional region space, ${(\interv{-1}{1}{})}^N$ is said to be
  the extended origin denoted $\textit{EO}_N$.
\end{defn}

\begin{defn}{(Origin distance)}
  We measure the \emph{origin distance} of a point offsets vector from the
  extended origin as the minimum number of translations by $\pm 1$ unit in one
  dimension at a time to reach any point in the extended origin. Since point
  offsets vector may represent infinite number of points, we measure the
  distance from the closest one.
\end{defn}
%
\begin{prop}{}
  The closed form of origin distance of an $N$ dimensional point offsets vector
  $p$ is the following:
%
  \begin{align*}
    \textit{dist}_i(p) & = \begin{cases}
      0 & \pi_i(p) = \interv{-\infty}{\infty}{} \\
      |\mathsf{a}| - 1 & \pi_i(p) = \{\mathsf{a}\} \\
    \end{cases} &
%
    |p| & = \sum_{i=1}^{N} dist_i(p)
  \end{align*}
\end{prop}
%
\begin{proof}
\end{proof}

>>>>

\begin{defn}{(\emph{Points})}
  We define the set \zinf{} of (one-dimensional) \emph{points} as
  $\mathbb{Z}$ extended with $\infty$ and $-\infty$. For any $n$ in
  \zinf{}, we have $-\infty \leq n \leq \infty$. The resulting set is
  a total order with top and bottom elements.
\end{defn}

\begin{defn}{(\emph{Point interval})}
  A \emph{point interval} is a closed interval $\zinf$ which may
  contain a ``hole'' at the origin, written \interv{a}{b}{c} where $a$ and $b$
  are drawn from \zinf{} with $a \leq 0 \leq b$, $a$ is $-\infty$ \emph{iff} $b$
  is $\infty$, and $c$ is drawn from $\mathbb{B} = \{ \mathsf{true}, \mathsf{false} \}$. We
  define it as follows:
\marginpar{We might add the restriction to $[-\infty, \infty]$ later,
  but okay for now.}
%
  \begin{equation*}
    \interv{a}{b}{c} \triangleq
      \{ n \;|\; a \leq n \leq b \wedge (\neg c \implies n \neq 0) \}
  \end{equation*}

@@

\begin{defn}{(\emph{Offsets and offset spaces})}
  An \emph{offset} is a Cartesian product (vector) of $\zinf$ values representing an
  $N$-dimensional point. The set of $N$-dimensional offsets is
  written $\zinf^N$.  An $N$-dimensional \emph{offset space} is a set of
  offsets, i.e. a member of $\mathcal{P}(\zinf^N)$.

An $N$-dimensional offset space can be \emph{projected} into the set
of points in the $i^{\text{th}}$ dimension by $\pi_i : \mathcal{P}(\zinf{})^{\mkern-19mu{N}} \to
  \mathcal{P}(\zinf{})$ where $i \in \mathsf{Fin}(N)$. We refer to this projected set as the \emph{$i^{th}$
    component of an offset space}. 
\end{defn}

\begin{proposition}
The set of $N$-dimensional offset spaces
$\mathcal{P}{(\zinf{}^{\mkern-19mu{N}})}$ trivially forms a
  distributive lattice under the superset relation.
\end{proposition}

\begin{remark}
Note that $\mathcal{P}(\zinf^N)$ is isomorphic to a sets of point
vectors $\mathcal{P}(\zinf)^N$. Therefore, projection $\pi$ on 
offset spaces is defined by lifting normal Cartesian product
projection to sets.
\end{remark}

%\begin{defn}{(\emph{Offsets vector}, $\mathcal{P}(\zinf^N)$)}
%  An $N$ dimensional \emph{offset space vector} is a Cartesian product of offset
%  spaces denoted with $\textit{offs}_1 \times \cdots \times
%  \textit{offs}_N$.
%  The function $\pi_i : \mathcal{P}(\zinf{})^{\mkern-19mu{N}} \to
%  \mathcal{P}(\zinf{})$ retrieves the $i^{th}$ offset space.%
%
%  The set of $N$ dimensional offsets vectors is
%  $\mathcal{P}{(\zinf{}^{\mkern-19mu{N}})}$ and they trivially form a
%  distributive lattice under the superset relation.
%\end{defn}

\begin{defn}{(\emph{Subscript space})}
  An $N$-dimensional offset space $\textit{off}$ is a \emph{subscript space} if it
  satisfies:
%
\begin{equation*}
\forall i \in \textsf{Fin}(N) . \quad 
\pi_i (\textit{off}) = \{p\} \wedge p \in \mathbb{Z} 
\;\;
\vee
\;\;
\pi_i (\textit{off}) = \interv{-\infty}{\infty}{\mathsf{true}}
\end{equation*}
%
That is, the $i^{th}$ component of the space is either a singleton 
in $\mathbb{Z}$ or the infinite interval.

Subscript spaces model array subscript expressions in our source
langauge where each subscripted expression is either a 
constant translation of an induction variable within a loop or some
other expression which is constant with respect to the loop e.g. \texttt{a(i + 1, 0)}.
\end{defn}

\begin{lem}{}\label{lem:vector-intersect}
  Intersection distributes over offset spaces. If $\mathit{off}$ and
  $\mathit{off'}$ are $N$-dimensional offset spaces, then
%
  \begin{equation*}
    \mathit{off} \cap \mathit{off'} =
      \prod_{i = 1}^{N} \pi_i(\mathit{off}) \cap \pi_i(\mathit{off'})
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    \mathit{off} \; \cap \; \mathit{off'} &
    = \{x \mid \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off}) \}
      \;\cap\;
      \{x \mid \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off'}) \} \\
    & = \{x \mid \bigwedge_{1 \leq i \leq N }
      (x_i \in \pi_i(\mathit{off}) \wedge x_i \in \pi_i(\mathit{off'})) \} \\
    & = \prod_{i = 1}^{N} \pi_i(\mathit{off}) \cap \pi_i(\mathit{off'})
  \end{align*}
\end{proof}

\begin{lem}{}\label{lem:vector-union}
  Let $\mathit{off}$ and $\mathit{off'}$ be two $N$ dimensional offset
  spaces such that $\pi_i(\mathit{off})$ is equal to $\pi_i(\mathit{off'})$ for all $1
  \leq i \leq N$ except possibly at $k$, then we have the following:
%
  \begin{equation*}
    \mathit{off} \; \cup \; \mathit{off'}
    =
    \pi_1(\mathit{off}) \times \cdots \times
    (\pi_k(\mathit{off}) \cup \pi_k(\mathit{off'})) \times \cdots \times
    \pi_N(\mathit{off})
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    \mathit{off} \; \cup \; \mathit{off'} &
    = \{x \mid
      \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off}) \}
      \;\cup\;
      \{x \mid
          \bigwedge_{1 \leq i \leq N } x_i \in \pi_i(\mathit{off'}) \} \\
    & = \{x \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}}
            x_i \in \pi_i(\mathit{off}) \wedge x_k \in \pi_k(\mathit{off}) \vee
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in
            \pi_i(\mathit{off'}) \wedge x_k \in \pi_k(\mathit{off'})
        \} \\
    & = \{x \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in
            \pi_i(\mathit{off}) \wedge
            x_k \in \pi_k(\mathit{off}) \cup \pi_k(\mathit{off'})
        \} \\
        & = \pi_1(\mathit{off}) \times \cdots \times
        (\pi_k(\mathit{off}) \cup \pi_k(\mathit{off'})) \times \cdots \times
        \pi_N(\mathit{off})
  \end{align*}
\end{proof}

\begin{defn}{(\emph{Offset interval})}
An offset interval is a finite Cartesian product of point intervals.
We referer to an interval vector of size $N$ as being $N$-dimensional, and
denoted by the set $\textit{Interval}^N$.
\end{defn}

\begin{defn}{(\emph{Region})} 
  A region \region{N} is a set of $N$-dimensional offset intervals defined as
  the smallest set satisfying the following:
%
  \begin{enumerate}
    \item If $R$ is in $\textit{Interval}^N$, then $R$ is in \region{N}.
    \item If $R$ and $S$ are in \region{N}, then so are $R \cap S$ and
      $R \cup S$.
  \end{enumerate}
\end{defn}

@@

Here helper function $\textit{peel} : \mathsf{Approx} \; a \to a$ removes the
approximation label.

@@

\begin{defn}{(Origin)}
In $N$-dimensions, the \emph{origin} $\mathbf{0}_N$ is the singleton set
$\{0^N\}$, or equivalently the vector of unit intervals $[0, 0]^N$.
\end{defn}

\begin{defn}{(Extended origin)}
In $N$-dimensions, the \emph{extended origin} is the
interval vector $[-1, 1]^N$ which is a subspace of \region{N}. 
\end{defn}

\begin{defn}{(Extended origin distance)}
  We measure the \emph{extended origin distance} of an offset from the
  extended origin as the minimum number of translations by $\pm 1$ in one
  dimension at a time, to reach an offset in the extended
  origin.

This notion extends to Since an 
  offsets vector may represent infinite number of points, we measure the
  distance from the closest one.
\end{defn}
%
\begin{prop}{}
  The closed form of origin distance of an $N$ dimensional point offsets vector
  $p$ is the following:
%
  \begin{align*}
    \textit{dist}_i(p) & = \begin{cases}
      0 & \pi_i(p) = \interv{-\infty}{\infty}{} \\
      |\mathsf{a}| - 1 & \pi_i(p) = \{\mathsf{a}\} \\
    \end{cases} &
%
    |p| & = \sum_{i=1}^{N} dist_i(p)
  \end{align*}
\end{prop}
%
\begin{proof}
\end{proof}
